Implementing Distributed Systems Directly\\with I/O Automata in ioa++

### 2011-12-14 papers/paper1/component_model.tex
Fixed typos.

We note that system action requests may fail for a variety of reasons.
A create fails if the automaton that should be added to the set of automata, i.e., the automaton being created, already exists.
A bind fails if the automaton associated with the output action or the automaton associated with the input action doesn't exist, or if any of the binding rules outlined in Section~\ref{component_structure} would be violated.
An unbind fails if the binding does not exist.
Note that the owner of the binding is taken into account when determining whether the binding exists, such that only the owner of a binding can dissolve it.
A destroy fails if the automaton being destroyed does not exist or is not a child of the automaton requesting the destroy action.

@@

% I didn't do more complex transactions, i.e., create-and-bind, because its not clear how it fails.
% Are partial successes allowed?
% Furthermore, such  transactions can be arbitrarily complex which doesn't bode well for efficiency.

>>>>

We note that system action requests may fail for a variety of reasons.
A create fails if the automaton that should be added to the set of automata, i.e., the automaton being created, already exists.
A bind fails if the automaton associated with the output action or the automaton associated with the input action does not exist, or if any of the binding rules outlined in Section~\ref{component_structure} would be violated.
An unbind fails if the binding does not exist.
Note that the owner of the binding is taken into account when determining whether the binding exists, such that only the owner of a binding can dissolve it.
A destroy fails if the automaton being destroyed does not exist or is not a child of the automaton requesting the destroy action.

@@

% I didn't do more complex transactions, i.e., create-and-bind, because its not clear how it fails.
% Are partial successes allowed?
% Furthermore, such  transactions can be arbitrarily complex which does not bode well for efficiency.

### 2011-12-14 papers/paper1/evaluation.tex
presubmission tweaks
Fixed typos.

The driver for a single trial can be found in examples/random.cpp\footnote{\url{github.com/jrwilson/ioa/blob/master/examples/random.cpp}} of the ioa++ package.
The trials were performed on a Mac G4 running OS X 10.5.8 with two 2.66 GHz Dual-Core Intel Xeon processors with 2GB of memory.
The code was compiled and linked with i686-apple-darwin9-g++-4.0.1 build 5493 with -O2 optimization and the -DPROFILE flag to enable profiling in ioa++.

@@

Figures~\ref{overhead1} and \ref{overhead2} show the average per thread of the fraction of time devoted to framework code and synchronization calls for the single and multi-threaded executions; confidence intervals are again negligible with largest being $\pm$0.002155 for Figure~\ref{overhead1} and $\pm$0.001781 for Figure~\ref{overhead2}.
Since the number of actions to be executed is constant, the single threaded execution shows a relatively consistent fraction for each value of $N$.
The fraction decreases as $\rho$ goes to 1 and the system transitions from acquiring two locks to one lock.
The multi-threaded execution shows that the fraction of time devoted to system overhead decreases as $\rho$ goes to 1 and is more pronounced when $N$ is large.
For example, when $N=1000000$ one thread spends half of its time waiting on the other thread when $\rho = 0$ and spends very little time on synchronization when $\rho = 1$.

>>>>

The driver for a single trial can be found in examples/random.cpp\footnote{\url{github.com/jrwilson/ioa/blob/master/examples/random.cpp}} of the ioa++ package.
The trials were performed on a Mac Pro running OS X 10.5.8 with two 2.66 GHz Dual-Core Intel Xeon processors with 2GB of memory.
The code was compiled and linked with i686-apple-darwin9-g++-4.0.1 build 5493 with -O2 optimization and the -DPROFILE flag to enable profiling in ioa++.

@@

Figures~\ref{overhead1} and \ref{overhead2} show the average per thread of the fraction of time devoted to framework code and synchronization calls for the single and multi-threaded executions; confidence intervals are again negligible with largest being $\pm$0.002155 for Figure~\ref{overhead1} and $\pm$0.001781 for Figure~\ref{overhead2}.
Since the number of actions to be executed is constant, the single threaded execution shows a relatively consistent fraction for each value of $N$.
The fraction decreases as $\rho$ goes from 0 to 1 because the system transitions from always acquiring two locks to always acquiring one lock.
The multi-threaded execution shows that the fraction of time devoted to system overhead decreases as $\rho$ goes to 1 and is more pronounced when $N$ is large.
For example, when $N=1000000$ one thread spends half of its time waiting on the other thread when $\rho = 0$ and spends very little time on synchronization when $\rho = 1$.

### 2011-12-14 papers/paper1/programming_model.tex
presubmission tweaks
Fixed typos.

\paragraph*{Declaring/defining actions}
Actions are member variables that model the \emph{action} 
\ifjournal
concept.
\else
concept\footnote{The term ``concept'' is used here in the sense of generic programming in C++, i.e., a collection of type requirements imposed by a C++ template.}.
\fi
Actions contain methods for preconditions, effects, and scheduling along with \emph{traits} that indicate the action's type (input, output, internal), value status, value type, parameter status, and parameter type.
These traits are used to guide the execution of the methods by the model and to supply values and parameters as necessary.
The traits are also used to check that input actions are bound to output actions with the same value status and type.
Bindings are checked for type compatibility at compile time using the generic programming facilities of C++.

@@

\paragraph*{Handling errors by replacement}
The non-deterministic and asynchronous nature of system actions complicates replacing an automaton.
To replace a child automaton, the parent automaton must destroy the child (thus dissolving all bindings with the child), create a new child, and bind to the new child.
Recall that system actions are asynchronous and can be executed in any order.
Thus, the user might might request the actions in one order, e.g., destroy, create, bind, but the system may execute them in a different order, e.g., create, bind, destroy.
If the sequence executed by the system automaton is (create, bind, destroy) then the bind will fail because the user automaton's actions are still bound to the old child.
Note that executing the bind action before the create will also fail but this case is prevented by the ioa::binding\_manager class.
Thus, a user automaton must receive notification that the first child has been destroyed before (creating and) binding to the second child.
A general solution is to observe the automaton and wait for it to be destroyed.

>>>>

\paragraph*{Declaring/defining actions}
Actions are member variables that model the \emph{action} 
\ifjournal
concept.
\else
concept\footnote{The term ``concept'' is used here in the sense of generic programming in C++, i.e., a collection of type requirements imposed by a C++ template.}.
\fi
Actions contain methods for preconditions, effects, and scheduling along with \emph{traits} that indicate the action's type (input, output, internal), value status, value type, parameter status, and parameter type.
These traits are used to guide the execution of the methods by the dispatcher and to supply values and parameters as necessary.
The traits are also used to check that input actions are bound to output actions with the same value status and type.
Bindings are checked for type compatibility at compile time using the generic programming facilities of C++.

@@

\paragraph*{Handling errors by replacement}
The non-deterministic and asynchronous nature of system actions complicates replacing an automaton.
To replace a child automaton, the parent automaton must destroy the child (thus dissolving all bindings with the child), create a new child, and bind to the new child.
Recall that system actions are asynchronous and can be executed in any order.
Thus, the user might request the actions in one order, e.g., destroy, create, bind, but the system may execute them in a different order, e.g., create, bind, destroy.
If the sequence executed by the system automaton is (create, bind, destroy) then the bind will fail because the user automaton's actions are still bound to the old child.
Note that executing the bind action before the create will also fail but this case is prevented by the ioa::binding\_manager class.
Thus, a user automaton must receive notification that the first child has been destroyed before (creating and) binding to the second child.
A general solution is to observe the automaton and wait for it to be destroyed.
